

library
```{r}
library(dplyr)
```



##
```{r}

df = readr::read_csv("~/Documents/projects/application_projects/20250115_NextSeq2000_2575_coverage_assembly/cov_GC/reports/20250115_NextSeq2000_cov_w_GC.csv")

#Bacillus_cereus_A2001_F06

df = df %>% 
  dplyr::mutate( reference = gsub( "(_[^_]*){2}$", "", sample_ID)) %>% 
  dplyr::mutate( well = gsub(".*_([A-Z0-9]+_[A-Z0-9]+)$", "\\1", sample_ID)) %>% 
  dplyr::group_by( well) %>% 
  dplyr::mutate( cov_mean = mean(cov)) %>% 
  dplyr::mutate( position = row_number()) %>% 
  dplyr::ungroup() %>% 
  dplyr::mutate( normalized_cov = cov/cov_mean) %>% 
  dplyr::mutate ( protocal = ifelse( grepl("(08|09)", sample_ID), "EP_std", "other")) %>% 
  dplyr::mutate ( protocal = ifelse( grepl("(06|07)", sample_ID), "EP_modified", protocal)) %>% 
  dplyr::mutate ( protocal = ifelse( grepl("(10|11)", sample_ID), "EPHC", protocal)) 


df

table(df$reference)

unique(df$reference)
```


Bacterium	GC Content (%)
Bacillus cereus	35
Bacillus subtilis	43
Clostridioides difficile	28
Enterobacter cloacae	55
Escherichia coli K-12	51
Pseudomonas aeruginosa	66
Rhodobacter sphaeroides	69
Staphylococcus epidermidis	32


Bacterium	GC_Content (%)
Bacillus_cereus	35
Bacillus_subtilis	43
Clostridioides_difficile	28
Enterobacter_cloacae	55
Escherichia_coli-K12_ATCC	51
Pseudomonas_aeruginosa	66
Rhodobacter_sphaeroides	69
Staphylococcus_epidermidis	32



```{r}

df_ref_gc = readr::read_table("GC_reference.csv")
names(df_ref_gc) = c("reference", "GC_content_genome")

df = df %>% 
  dplyr::left_join( df_ref_gc)
  
names(df)
```



```{r}

df_by_group = df %>% 
  dplyr::mutate( group = ifelse (grepl("(09|07|11)", sample_ID ),  "rep2", "rep1") ) %>% 
  dplyr::group_by( sample_ID) %>% 
  dplyr::mutate( avg_gc_content = mean(GC_content, na.rm =T)) %>% 
  dplyr::ungroup() %>% 
  dplyr::mutate( gc_diff = GC_content - avg_gc_content,
                 well = paste0(well, "_", protocal))
  

table(df_by_group$avg_gc_content)
  

table(df_by_group$group)

df_group_1 = df_by_group %>% 
  dplyr::filter( group == "rep1")

df_group_2 = df_by_group %>% 
  dplyr::filter( group == "rep2")

df_group_1
```



```{r}

df_test_ref = df_group_1 %>%
    dplyr::filter( reference ==  "Bacillus_cereus") 

summary(df_test_ref$gc_diff)


df_group_1 %>%
    dplyr::filter(reference == "Bacillus_cereus") %>%
    ggplot(aes(x = position, y = normalized_cov, fill = gc_diff)) +
    geom_bar(stat = "identity") +  # Ensures the area follows the actual values
    facet_wrap(~ protocal, scales = "free", ncol = 3) +
    xlab("Position in kb") +
    scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = mean(df_group_1$gc_diff, na.rm = TRUE)) +
    theme(strip.text = element_text(size = 8))





```

## for group 1

```{r}

library(ggplot2)
library(dplyr)
library(ggpubr)

# Modify plot_by_genome to remove x-axis labels but keep x-axis ticks
plot_by_genome = function(ref){
  plot = df_group_1 %>%
    dplyr::mutate( protocal = paste0( ref, " GC ",  GC_content_genome , ": ", protocal )) %>% 
    dplyr::filter(reference == ref) %>%
    ggplot(aes(x = position, y = normalized_cov, fill = gc_diff)) +
    geom_bar(stat = "identity") +
    facet_wrap(~ protocal, scales = "free", ncol = 3) +
    xlab("Position in kb") +  # This will be removed later except for the last plot
    scale_fill_gradient2(low = "blue", mid = "white", high = "red", 
                         midpoint = mean(df_group_1$gc_diff, na.rm = TRUE)) +
    theme(strip.text = element_text(size = 8), 
          axis.title.x = element_blank())  # Remove only the x-axis label
  
  return(plot)
}

# List of genome references
genomes <- c("Bacillus_cereus", "Bacillus_subtilis", "Clostridioides_difficile", 
             "Enterobacter_cloacae", "Escherichia_coli-K12_ATCC", 
             "Pseudomonas_aeruginosa", "Rhodobacter_sphaeroides", "Staphylococcus_epidermidis")

# Generate plots
plots <- lapply(genomes, plot_by_genome)

# Add back x-axis label only for the last plot
plots[[length(plots)]] <- plots[[length(plots)]] +
  xlab("Position in kb")  # Restore the x-axis label for the last plot only

# Arrange the plots
combined_plot <- ggarrange(plotlist = plots, ncol = 1, nrow = 8, align = "v")

# Save the figure
ggsave("summary_figures/rep1_20250115_Nextseq2000_normalized_cov_GC.jpg", 
       plot = combined_plot, width = 12, height = 16, dpi = 300)

```







## for group 2

```{r}

library(ggplot2)
library(dplyr)
library(ggpubr)

# Modify plot_by_genome to remove x-axis labels but keep x-axis ticks
plot_by_genome = function(ref){
  plot = df_group_2 %>%
    dplyr::mutate( protocal = paste0( ref, " GC ",  GC_content_genome , ": ", protocal )) %>% 
    dplyr::filter(reference == ref) %>%
    ggplot(aes(x = position, y = normalized_cov, fill = gc_diff)) +
    geom_bar(stat = "identity") +
    facet_wrap(~ protocal, scales = "free", ncol = 3) +
    xlab("Position in kb") +  # This will be removed later except for the last plot
    scale_fill_gradient2(low = "blue", mid = "white", high = "red", 
                         midpoint = mean(df_group_1$gc_diff, na.rm = TRUE)) +
    theme(strip.text = element_text(size = 8), 
          axis.title.x = element_blank())  # Remove only the x-axis label
  
  return(plot)
}



# Generate plots
plots <- lapply(genomes, plot_by_genome)

# Add back x-axis label only for the last plot
plots[[length(plots)]] <- plots[[length(plots)]] +
  xlab("Position in kb")  # Restore the x-axis label for the last plot only

# Arrange the plots
combined_plot <- ggarrange(plotlist = plots, ncol = 1, nrow = 8, align = "v")

# Save the figure
ggsave("summary_figures/rep2_20250115_Nextseq2000_normalized_cov_GC.jpg", 
       plot = combined_plot, width = 12, height = 16, dpi = 300)

```

